---
title: "STAT547 Homework #10"
output: github_document
---
### Make API queries "by hand" using httr 

Load the following packages:
```{r results='hide', message=FALSE, warning=FALSE}
library(httr)
library(purrr)
library(glue)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(plyr)
library(knitr)
library(forcats)
```


### Call an API from OpenWeatherData
Sign up [here](https://home.openweathermap.org/users/sign_up) to request an API from OpenWeatherData. Please note it will take 10 minutes for the activation of an API key.
```{r}
my_apikey <- "insert_your_api_key_here"

#Current weather in several Canadian cities
weather_severalcities <- GET(glue("http://api.openweathermap.org/data/2.5/group?id=6173331,6167865,6077243,5969423,5913490,6183235,6094817,5969785&units=metric&APPID={my_apikey}")) 
#changed the units from the default which is Kelvin to metric which is Celsius
#`GET` is used to retrieve resources

#You must input several city ID's into `weather_severalcities` in order to get weather data for the specific cities you want. The API query above includes data for the following cities:
vancouver_id <- "6173331" #Vancouver,CA
toronto_id <- "6167865" #Toronto,CA
montreal_id <- "6077243" #Montreal,CA
halifax_id <- "5969423" #Halifax,CA
calgary_id <- "5913490" #Calgary,CA
winnipeg_id <- "6183235" #Winnipeg,CA
ottawa_id <- "6094817" #Ottawa,CA
hamilton_id <- "5969785" #Hamilton,CA
#A full list of city ID (city.list.json.gz) can be downloaded here http://bulk.openweathermap.org/sample/


status_code(weather_severalcities) #status code of 200 means that the call was successful 

severalcities_result <- content(weather_severalcities)
nchar(severalcities_result) #number of characters
class(severalcities_result) #this is a list
length(severalcities_result) #the list is of length 2

View(severalcities_result)

str(weather_severalcities)
```

List of data frames for the different cities.
```{r}
#data frame for the weather data from Vancouver
weather_df_1 <- lapply(severalcities_result, function(x){
  df <- data_frame(City = severalcities_result$list[[1]]$name,
                   Longitude = severalcities_result$list[[1]]$coord$lon,
                   Latitude = severalcities_result$list[[1]]$coord$lat,
                   Temperature = severalcities_result$list[[1]]$main$temp,
                   Pressure = severalcities_result$list[[1]]$main$pressure,
                   Humidity = severalcities_result$list[[1]]$main$humidity,
                   Temp_min = severalcities_result$list[[1]]$main$temp_min,
                   Temp_max = severalcities_result$list[[1]]$main$temp_max,
                   Wind_speed = severalcities_result$list[[1]]$wind$speed)
}) %>% bind_rows()

#data frame for the weather data from Toronto
weather_df_2 <- lapply(severalcities_result, function(x){
  df <- data_frame(City = severalcities_result$list[[2]]$name,
                   Longitude = severalcities_result$list[[2]]$coord$lon,
                   Latitude = severalcities_result$list[[2]]$coord$lat,
                   Temperature = severalcities_result$list[[2]]$main$temp,
                   Pressure = severalcities_result$list[[2]]$main$pressure,
                   Humidity = severalcities_result$list[[2]]$main$humidity,
                   Temp_min = severalcities_result$list[[2]]$main$temp_min,
                   Temp_max = severalcities_result$list[[2]]$main$temp_max,
                   Wind_speed = severalcities_result$list[[2]]$wind$speed)
}) %>% bind_rows()

#data frame for the weather data from Montreal
weather_df_3 <- lapply(severalcities_result, function(x){
  df <- data_frame(City = severalcities_result$list[[3]]$name,
                   Longitude = severalcities_result$list[[3]]$coord$lon,
                   Latitude = severalcities_result$list[[3]]$coord$lat,
                   Temperature = severalcities_result$list[[3]]$main$temp,
                   Pressure = severalcities_result$list[[3]]$main$pressure,
                   Humidity = severalcities_result$list[[3]]$main$humidity,
                   Temp_min = severalcities_result$list[[3]]$main$temp_min,
                   Temp_max = severalcities_result$list[[3]]$main$temp_max,
                   Wind_speed = severalcities_result$list[[3]]$wind$speed)
}) %>% 
  bind_rows()

#data frame for the weather data from Halifax
weather_df_4 <- lapply(severalcities_result, function(x){
  df <- data_frame(City = severalcities_result$list[[4]]$name,
                   Longitude = severalcities_result$list[[4]]$coord$lon,
                   Latitude = severalcities_result$list[[4]]$coord$lat,
                   Temperature = severalcities_result$list[[4]]$main$temp,
                   Pressure = severalcities_result$list[[4]]$main$pressure,
                   Humidity = severalcities_result$list[[4]]$main$humidity,
                   Temp_min = severalcities_result$list[[4]]$main$temp_min,
                   Temp_max = severalcities_result$list[[4]]$main$temp_max,
                   Wind_speed = severalcities_result$list[[4]]$wind$speed)
}) %>% bind_rows()

#data frame for the weather data from Calgary
weather_df_5 <- lapply(severalcities_result, function(x){
  df <- data_frame(City = severalcities_result$list[[5]]$name,
                   Longitude = severalcities_result$list[[5]]$coord$lon,
                   Latitude = severalcities_result$list[[5]]$coord$lat,
                   Temperature = severalcities_result$list[[5]]$main$temp,
                   Pressure = severalcities_result$list[[5]]$main$pressure,
                   Humidity = severalcities_result$list[[5]]$main$humidity,
                   Temp_min = severalcities_result$list[[5]]$main$temp_min,
                   Temp_max = severalcities_result$list[[5]]$main$temp_max,
                   Wind_speed = severalcities_result$list[[5]]$wind$speed)
}) %>% bind_rows()

#data frame for the weather data from Winnipeg
weather_df_6 <- lapply(severalcities_result, function(x){
  df <- data_frame(City = severalcities_result$list[[6]]$name,
                   Longitude = severalcities_result$list[[6]]$coord$lon,
                   Latitude = severalcities_result$list[[6]]$coord$lat,
                   Temperature = severalcities_result$list[[6]]$main$temp,
                   Pressure = severalcities_result$list[[6]]$main$pressure,
                   Humidity = severalcities_result$list[[6]]$main$humidity,
                   Temp_min = severalcities_result$list[[6]]$main$temp_min,
                   Temp_max = severalcities_result$list[[6]]$main$temp_max,
                   Wind_speed = severalcities_result$list[[6]]$wind$speed)
}) %>% bind_rows()

#data frame for the weather data from Ottawa
weather_df_7 <- lapply(severalcities_result, function(x){
  df <- data_frame(City = severalcities_result$list[[7]]$name,
                   Longitude = severalcities_result$list[[7]]$coord$lon,
                   Latitude = severalcities_result$list[[7]]$coord$lat,
                   Temperature = severalcities_result$list[[7]]$main$temp,
                   Pressure = severalcities_result$list[[7]]$main$pressure,
                   Humidity = severalcities_result$list[[7]]$main$humidity,
                   Temp_min = severalcities_result$list[[7]]$main$temp_min,
                   Temp_max = severalcities_result$list[[7]]$main$temp_max,
                   Wind_speed = severalcities_result$list[[7]]$wind$speed)
}) %>% bind_rows()

#data frame for the weather data from Hamilton
weather_df_8 <- lapply(severalcities_result, function(x){
  df <- data_frame(City = severalcities_result$list[[8]]$name,
                   Longitude = severalcities_result$list[[8]]$coord$lon,
                   Latitude = severalcities_result$list[[8]]$coord$lat,
                   Temperature = severalcities_result$list[[8]]$main$temp,
                   Pressure = severalcities_result$list[[8]]$main$pressure,
                   Humidity = severalcities_result$list[[8]]$main$humidity,
                   Temp_min = severalcities_result$list[[8]]$main$temp_min,
                   Temp_max = severalcities_result$list[[8]]$main$temp_max,
                   Wind_speed = severalcities_result$list[[8]]$wind$speed)
}) %>% bind_rows()
```

Now I must recursively join the list of data frames:

*I found [this](https://www.rdocumentation.org/packages/plyr/versions/1.8.4/topics/join_all) website particularly helpful with figuring out how to join multiple data frames 
```{r}
weather_cities_df <- join_all(list(
  weather_df_1, 
  weather_df_2, 
  weather_df_3, 
  weather_df_4, 
  weather_df_5, 
  weather_df_6, 
  weather_df_7, 
  weather_df_8),
  by = "City", type = "full", match = "all")

#Now I need to remove duplicates from the `weather_cities_df` data frame using the function `distinct()`
weather_cities_df <- distinct(weather_cities_df,
                              City,
                              Longitude,
                              Latitude,
                              Temperature,
                              Pressure,
                              Humidity,
                              Temp_min,
                              Temp_max,
                              Wind_speed)
kable(weather_cities_df, align = 'c', format = 'html')
```

Now to test to be sure the data has been successfully downloaded and cleaned.
```{r}
#Plot of temperatures for the different cities
(weather_cities_plot <- weather_cities_df %>%
  ggplot(aes(x = City,
             y = Temperature)) +
  geom_point() +
   theme_bw()+
   theme(axis.title = element_text(size=13, face='bold'),
         axis.text.x = element_text(size=12, angle=90),
        axis.text.y = element_text(size=12),
        plot.title = element_text(size=14, face="bold", hjust = 0.5)) +
   labs(x = "Cities in Canada", y = "Temperature (?C)", title = "Temperature for cities on December 7, 2017", caption = "Data from OpenWeatherMap"))

#Determine which cities have a temperature above zero degrees Celsius and a wind speed less than 5.0 meters/sec
(test1 <- weather_cities_df %>%
  filter(Temperature > 0, Wind_speed <5.0) %>%
  group_by(City) %>%
    kable())
#as you can see only Vancouver and Calgary fit the requirements
```


