---
title: "STAT 547 Homework Assignment #6"
output: github_document
---

[Return to Homework 6 Folder](https://github.com/heathersummers/STAT547-hw-Summers-Heather/tree/master/hw06)

[Return to Main Page of Repository](https://github.com/heathersummers/STAT547-hw-Summers-Heather)

## Table of Contents:

- [Working with a nested data frame](#working-with-a-nested-data-frame)

    - [Linear Model](#linear-model)

    - [Quadratic Model](#quadratic-model)

- [Writing Functions](#writing-functions)



Load the following packages:
```{r results='hide', message=FALSE, warning=FALSE}
library(gapminder)
library(tidyverse)
library(broom)
library(forcats)
library(purrr)
library(robustbase)
```

## Task 6: 
### Working with a nested data frame
#### Linear Model

<a href="#top">Back to top</a>

Creating a nested data frame and fitting a linear model of life expectancy against year.
First, nest the data by country and continent.
```{r}
#Nesting the data by country and continent:
nested_gapminder<- gapminder %>%
  group_by(continent, country) %>% #keeps country and continent on the outside
  nest()
head(nested_gapminder)
```


```{r}
linear_fit <- function(dat, offset = 1952){
  lm(lifeExp ~ I(year - offset), data = dat)
}

linear_fit(nested_gapminder[[1, "data"]]) #model is fit to the data from Afghanistan

(fits <- map(nested_gapminder$data[1:2], linear_fit))

#`map()` inside `mutate()` in order to store the models inside the nested gapminder data set (`nested_gapminder`) in another list-column
(gapminder_coefficients <- nested_gapminder %>%
    mutate(fit = map(data, linear_fit),
           clean = map(fit, tidy)) %>%
  select(continent, country, clean) %>% 
  unnest(clean)) #unnest is used to simplify back to a normal tibble
```

Exploring the fit of the model using functions from the `broom` package:
```{r message=FALSE, warning=FALSE}
#Tidy the data for `lm` objects produces a data frame summary of estimated coefficients and results related to statistical inference such as the p-value.
gfits_broom <- gapminder %>%
  group_by(country, continent) %>% 
  do(tidy(lm(lifeExp ~ I(year - 1952), data = .))) #`tidy()` is a tidy version of `summary()` and gives a table with one row per parameter estimate
gfits_broom


#Can use several other functions from the `broom` package but it is more convient to first store the fitted model as an object (ex. `fits`).
(fits <- gapminder %>%
  group_by(country, continent) %>% 
  do(fit = lm(lifeExp ~ I(year - 1952), .)))

fits %>%
  glance(fit) #one row per country and it provides statistics on the overall model inclduing the R squared value and adjusted R squared value.

fits %>% 
  augment(fit) #one row per original observation and it gives several summary statistics including fitted value, residual, Cook's D, etc.
```

Exploring the R squared values of the linear model:

Analyzed which countries had an R squared value less than 0.8 when fitting the linear model of life expectancy against the year 1952. 
```{r}
(fits %>% 
  glance(fit) %>%
  filter(r.squared<0.8) %>% 
  ggplot(aes(x=r.squared, y=country, colour = continent)) +
  geom_point() +
  scale_x_continuous ("R Squared Values of Linear Model") +
  labs(title = "R Squared Values for Linear Model Based on 1952 Life Expectancy Data", y = "Country", colour = "Continent") +
   theme_bw()+
   theme(axis.title = element_text(size = 12, face ="bold"),
         axis.text.x = element_text(size = 10),
         axis.text.y = element_text(size = 8),
         plot.title = element_text(size = 13, face = "bold", hjust = 0.5),
         legend.title = element_text(size = 10, face = "bold")))

```


Exploring the estimated coefficients:
```{r}
(estimate_coef <- gapminder_coefficients %>%
   mutate(country = fct_reorder(country, estimate)) %>%
   filter(continent=="Asia", (estimate>1)) %>%
   ggplot(aes(x = estimate, y = country)) +
   geom_point() +
   scale_x_continuous ("Life Expectancy Estimated Coefficients") +
   labs(title = "Estimated Coefficients Based on Linear Model for Asia in 1952", y = "Country") +
   theme_bw()+
   theme(axis.title = element_text(size = 12, face ="bold"),
         axis.text.x = element_text(size = 10),
         axis.text.y = element_text(size = 8),
         plot.title = element_text(size = 13, face = "bold", hjust = 0.5)))
```

#### Quadratic Model

<a href="#top">Back to top</a>

Using the nested data frame and fitting a quadratic model of life expectancy against year.
```{r}
quad_fit <- function(dat, offset = 1952){
  lm(lifeExp ~ I(year - offset) + I(year^2), data = dat)
}
quad_fit(nested_gapminder[[1, "data"]]) #model is fit to the data from Afghanistan

(quadfits <- map(nested_gapminder$data[1:2], quad_fit))

(gapminder_coefficients <- nested_gapminder %>%
    mutate(fitted = map(data, quad_fit),
           tidy = map(fitted, tidy)) %>%
  select(continent, country, tidy) %>%
  unnest(tidy))
```

```{r message=FALSE, warning=FALSE}
(quadfits <- gapminder %>%
  group_by(country, continent) %>% 
  do(fit2 = lm(lifeExp ~ I(year - 1952)+ I(year^2), .)))

quadfits %>%
  glance(fit2) #one row per country and it provides statistics on the overall model inclduing the R squared value and adjusted R squared value.

quadfits %>% 
  augment(fit2) #one row per original observation and it gives several summary statistics including fitted value, residual, Cook's D, etc.
```

Exploring the R squared values of the quadratic model:

Analyzed which countries had an R squared value less than 0.8 when fitting the quadratic model of life expectancy against the year 1952. 
```{r}
(quadfits %>% 
  glance(fit2) %>%
  filter(r.squared<0.8) %>% 
  ggplot(aes(x=r.squared, y=country, colour = continent)) +
  geom_point() +
  scale_x_continuous ("R Squared Values of Quadratic Model") +
  labs(title = "R Squared Values for Quadratic Model Based on 1952 Life Expectancy Data", y = "Country", colour = "Continent") +
   theme_bw()+
   theme(axis.title = element_text(size = 12, face ="bold"),
         axis.text.x = element_text(size = 10),
         axis.text.y = element_text(size = 8),
         plot.title = element_text(size = 13, face = "bold", hjust = 0.5),
         legend.title = element_text(size = 10, face = "bold")))

```

Based on the figures of R squared values, it appears the quadratic model is a better fit of the nested data (i.e. less countries than have an R squared value less than 0.8).

Exploring the estimated coefficients:
```{r}
(estimate_coef <- gapminder_coefficients %>%
   mutate(country = fct_reorder(country, estimate)) %>%
   filter(continent=="Asia", (estimate>1)) %>%
   ggplot(aes(x = estimate, y = country)) +
   geom_line() +
   scale_x_continuous ("Life Expectancy Estimated Coefficients") +
   labs(title = "Estimated Coefficients Based on Quadratic Model for Asia in 1952", y = "Country") +
   theme_bw()+
   theme(axis.title = element_text(size = 12, face ="bold"),
         axis.text.x = element_text(size = 10),
         axis.text.y = element_text(size = 8),
         plot.title = element_text(size = 13, face = "bold", hjust = 0.5)))
```

## Task 2: 
### Writing Functions

<a href="#top">Back to top</a>

The goal of this function is for a linear regression function to be generalized to do a quadratic regresion and a robust regression (`lmrob()`).

First, I nested the data by country.
```{r}
gap_nested <- gapminder %>%
  group_by(country) %>% #keeps country on the outside
  nest()
head(gap_nested)
```

If robust is used than `robust` will be equal to 1 and if it is not used it will be equal to 0. If it is a linear regression function than `model` will be equal to 1 and if it is a quadratic regression function than `model` will be equal to 2. 

```{r message=FALSE, warning=FALSE}
#linear function: lm(gdpPercap ~ I(year - offset)
#quadratic function: lm(gdpPercap ~ I(year - offset) + I(year^2)

fit_model <- function(data, offset = 2007, robust = 1, model = 2){
  if(robust==1) lmrob(gdpPercap ~ I(year - offset) + I(year^2*model), data = data, control = lmrob.control(max.it = 100)) else
    lm(gdpPercap ~ I(year - offset) + I(year^2*model), data = data)
}

#lin_fit_norob = linear fit without robust
#lin_fit_rob = linear fit with robust
#quad_fit_norob = quadratic fit without robust
#quad_fit_rob = quadratic fit with robust

gap_nested <- gap_nested %>%
  mutate(lin_fit_norob = map(data, fit_model, 0, 1), #map is used for doing computations involving list columns
         lin_fit_rob = map(data, fit_model, 1, 2),
         quad_fit_norob = map(data, fit_model, 0, 1),
         quad_fit_rob = map(data, fit_model, 1, 2))
head(gap_nested)
```


## Report Your Process:

<a href="#top">Back to top</a>

I first started doing task #3 but struggled with understanding the `stringr` package including how to clean the "non-word" text using various characters. I used the notes from past STAT545 classes but still found it difficult so I have left that for another time. I struggled with finding notes and websites that were helpful for writing functions but did find several useful links when I was working with the nested data frame (see below). Overall, I found this assignment much harder and more time consuming than any of the previous assignments. 


Useful Links:

[Fitting a Quadratic Model](http://www.theanalysisfactor.com/r-tutorial-4/)

[Computing by groups within data.frames with dplyr and broom](http://stat545.com/block023_dplyr-do.html)

[broom package](https://github.com/tidyverse/broom)

[Split-Apply-Combine](http://stat545.com/block024_group-nest-split-map.html)

[Robust regression](https://stats.idre.ucla.edu/r/dae/robust-regression/)

[Linear regression models with robust parameter estimation](https://www.r-bloggers.com/linear-regression-models-with-robust-parameter-estimation/)


<a href="#top">Back to top</a>